---
title: "p8105_hw3_tl3195"
author: "Tingyi Li"
date: "2023-10-13"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)

library(p8105.datasets)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# Problem 1 (Instacart)

Read in the dataset:

```{r}
data("instacart")
instacart=instacart
```

Description of dataset:
This dataset has `r nrow(instacart)` rows and `r ncol(instacart)` columns,  with each row resprenting a single product from an instacart order. 
Variables include:
`order_id`: order identifier
`product_id`: product identifier
`add_to_cart_order`: the order in which each product was added to the cart
`reordered`: 1 if this prodcut has been ordered by this user in the past, 0 otherwise
`user_id`: customer identifier
`eval_set`: which evaluation set this order belongs in (Note that the data for use in this class is exclusively from the “train” eval_set)
`order_number`: the order sequence number for this user (1=first, n=nth)
`order_dow`: the day of the week on which the order was placed
`order_hour_of_day`: the hour of the day on which the order was placed
`days_since_prior_order`: days since the last order, capped at 30, NA if order_number=1
product_name: name of the product
`aisle_id`: aisle identifier
`department_id`: department identifier
`aisle`: the name of the aisle
`department`: the name of the department

The number of aisles and which aisles are the most items ordered from:
```{r}
instacart |> 
  count(aisle) |> 
  arrange(desc(n))
```
There are `r n_distinct(instacart$aisle)` aisles. The most item are ordered from fresh vegetables aisles. 

Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered

```{r}
instacart |> 
  count(aisle) |> 
  filter(n > 10000) |> 
  mutate(aisle = fct_reorder(aisle, n)) |> 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

```

Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r}
instacart |>
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) |>
  group_by(aisle) |>
  count(product_name) |> 
  mutate(rank = min_rank(desc(n))) |> 
  filter(rank < 4) |> 
  arrange(desc(n)) |>
  knitr::kable()  
```

Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week.

```{r}
instacart |>
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) |>
  group_by(product_name, order_dow) |>
  summarize(mean_hour = mean(order_hour_of_day)) |>
  pivot_wider(
    names_from = order_dow, 
    values_from = mean_hour) |>
  knitr::kable(digits = 2)
```

# Problem 2 (BRFSS)

Load the data data from the p8105.datasets package using:
```{r}
data("brfss_smart2010")
brfss=brfss_smart2010
```
 
Data cleaning:

format the data to use appropriate variable names;
focus on the “Overall Health” topic
include only responses from “Excellent” to “Poor”
organize responses as a factor taking levels ordered from “Poor” to “Excellent”

Specify response column (show all the values):
```{r}
brfss=brfss |>
  janitor::clean_names() |>
  mutate(data_value=as.double(data_value))
unique(brfss$response)
```
The responses include "Excellent", "Very Good", "Good", "Fair", and "Poor".

Organize responses as a factor taking levels ordered from "Poor" to "Excellent":
```{r}
brfss=brfss|>
  filter(topic=="Overall Health",response %in% c("Poor", "Fair", "Good", "Very good", "Excellent"))|>
  mutate(
    response=factor(response, levels=c("Poor", "Fair", "Good", "Very good", "Excellent"))
  )
```
Therefore, the order now is "Poor", "Fair", "Good", "Very Good", and "Excellent".

In 2002, states were observed at 7 or more locations
```{r}
brfss |>
  filter(year==2002) |>
  group_by(locationabbr) |>
  summarize(n_stats = n_distinct(locationdesc)) |>
  filter(n_stats>=7)
```
CT, FL, MA, NC, NJ, and PA are states observed at 7 or more locations in 2002.

In 2010, states were observed at 7 or more locations
```{r}
brfss |>
  filter(year==2010) |>
  group_by(locationabbr) |>
  summarize(n_stats = n_distinct(locationdesc)) |>
  filter(n_stats>=7)
```
CA, CO, FL, MA, MD, NC, NE, NJ, NY, OH, PA, SC, TX, and WA are states observed at 7 or more locations in 2010.

Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state.
```{r}
dataone = brfss |>
  filter(response=="Excellent") |>
  group_by(year, locationabbr) |>
  summarize(data_value_mean=mean(data_value, na.rm = TRUE))
```

Make a “spaghetti” plot of this average value over time within a state:
```{r}
dataone |>
  rename("state"="locationabbr") |>
  ggplot(aes(x=year, y=data_value_mean, color=state)) +
  geom_point() + geom_line() +
  theme(legend.position = "bottom")
```

The average data_value in excellent responses is between year of 2002 to year of 2010. For most of the states, the range of the average data_value for excellent responses is approximately between 17 and 28.

Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.
```{r}
two_panel=
  brfss |>
  filter(locationabbr=="NY"& year %in% c(2006,2010))|>
  ggplot(aes(x =response , y = data_value)) + 
  geom_boxplot()+
  labs(title = "Distribution of data_value in NY State in 2006 and 2010") +
  facet_grid(.~year)

two_panel
```

In both 2006 and 2010 distributions, the "Poor" responses have lowest median value. Generally, the value of "Excellent" responses, "Very Good" responses and "Good" resposnes have higher data values than "Poor" and "Fair" responses in both 2006 and 2010. The "Very Good" responses in 2010 have larger variation and higher median value compared to those in 2006. The "Good" responses in 2010 have smaller variation lower median compare to those  in 2006. The "Fair" responses in 2010 have larger variations compared to those in 2006.

# Problem 3 (Accelerometers)

Import dataset
```{r}
demographic = read_csv("./nhanes_covar.csv", skip=4) |>
  janitor::clean_names()
accelerometer=read_csv("./nhanes_accel.csv") |>
  janitor::clean_names()
  
```

Clean the demograhic data:

Include all originally observed variables; 
Exclude participants less than 21 years of age, and those with missing demographic data;
Encode data with reasonable variable classes (i.e. not numeric, and using factors with the ordering of tables and plots in mind).
```{r}
demographic=demographic|>
  drop_na()|>
  filter(age>=21)|>
  mutate(
     sex=case_match(
       sex,
       1~"male",
       2~"female"),
     education=case_match(
       education,
       1~"Less than high school",
       2~"High school equivalent",
       3~"More than high school"
     ),
     education=factor(education, levels=c(
       "Less than high school",
       "High school equivalent",
       "More than high school"))
  )
```

Clean accelerometer data:
```{r}
accelerometer=accelerometer|>
  drop_na()|>
  pivot_longer(
    min1:min1440,
    names_to = "time",
    values_to = "mims_value"
  ) |>
  separate(time, into=c("min","minute"),3)|>
  select(-min)|>
  mutate(minute=as.integer(minute))
  
```

Merge the demographic data and accelerometer data:
```{r}
demo_acce=
  inner_join(demographic, accelerometer, by="seqn")
```

Produce a reader-friendly table for the number of men and women in each education category:
```{r}
demo_acce|>
  group_by(sex,education)|>
  summarize(n_obs=n_distinct(seqn))|>
  pivot_wider(
    names_from=education,
    values_from=n_obs
  )|>
  knitr::kable()
```
For both males and females, "More than high school" category has the highest observations. Therefore, the education level of both males and females are beyond high school. In "Less than high school" category, the number of males and the number of females are approximately the same (men is 27 and women is 28). In "High school equivalent" category, there are more males (35) than females (23).

Create a visualization of the age distributions for men and women in each education category:
```{r}
demo_acce|>
  group_by(education,seqn)|>
  distinct(sex,age,education,seqn)|>
  ggplot(aes(x=age, fill=education)) +
  geom_density(alpha = .5) +
  labs(title="Age Distribution") +
  facet_grid(.~sex)
```

For both males and females, the age distribution of "More than high school" category is right-skewed.
For both males and females, the age distribution of "less than high school" category is left-skewed. All these indicate that young people tend to have higher education level. 
There is a greater difference among age distribution in females than that in males, which means that gender influences the education level across age.

Create a total activity variable for each participant
```{r}
tot_act=demo_acce|>
  group_by(sex,age,education,bmi,seqn)|>
  summarise(total_activity=sum(mims_value))
```

Plot these total activities (y-axis) against age (x-axis);compare men to women and have separate panels for each education level
```{r}
tot_act|>
  ggplot(aes(x=age, y=total_activity, color=sex)) +
  geom_smooth() + geom_point(alpha=.75) +
  facet_grid(.~education)
```

For "High School Equivalent" and "More than High School" categories, the total activities for female is higher than the total activities for males. Both females and males at young age have larger total activties in all three education groups. For "More than High School" category, both females and males have a smaller variation in age than other two education groups.

Make a three-panel plot that shows the 24-hour activity time courses for each education level and use color to indicate sex.
```{r}
demo_acce|>
  group_by(sex,education,minute)|>
  summarise(mims_value=mean(mims_value))|>
  ggplot(aes(x=minute, y=mims_value, color=sex)) +
  geom_smooth() + geom_point(alpha=.75) +
  labs(title="24-hour activity time courses for each education level") +
  facet_grid(.~education)
```

For all three education levels, there seems to be a similar trend where the MIMS value are higher roughly around the 8:30 and 18:30 (500 to 1100 minute) for both females and males. MIMS value start to increase drastically at 4:17 and reaches the peak at 8:30. Then between 8:30 and 18:30, the MIMS value decreases slightly. After 18:30, MIMS value decreases in a substantial trend in all three education levels. For both "High school equivalent" category and "More than high school" category, females seem to have higher MIMS value compared to males between 8:30 and 18:30 since the line for females lies above the line for males. However, for "Less than high school" category, females and males seem to have similar MIMS value between 8:30 and 18:30. It's also apparent that both females and males in "Less than high school" category have higher MIMS values around the peak compared to the other two education levels. This indicates that education level seems to be correlated with the different MIMS values around the peak that lower education level is associated with the higher MIMS values around the peak. 