---
title: "p8105_hw3_tl3195"
author: "Tingyi Li"
date: "2023-10-13"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)

library(p8105.datasets)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# Problem 1 (Instacart)

Read in the dataset:

```{r}
data("instacart")
instacart=instacart
```

Description of dataset:
This dataset has `r nrow(instacart)` rows and `r ncol(instacart)` columns,  with each row resprenting a single product from an instacart order. 
Variables include:
`order_id`: order identifier
`product_id`: product identifier
`add_to_cart_order`: the order in which each product was added to the cart
`reordered`: 1 if this prodcut has been ordered by this user in the past, 0 otherwise
`user_id`: customer identifier
`eval_set`: which evaluation set this order belongs in (Note that the data for use in this class is exclusively from the “train” eval_set)
`order_number`: the order sequence number for this user (1=first, n=nth)
`order_dow`: the day of the week on which the order was placed
`order_hour_of_day`: the hour of the day on which the order was placed
`days_since_prior_order`: days since the last order, capped at 30, NA if order_number=1
product_name: name of the product
`aisle_id`: aisle identifier
`department_id`: department identifier
`aisle`: the name of the aisle
`department`: the name of the department

The number of aisles and which aisles are the most items ordered from:
```{r}
instacart |> 
  count(aisle) |> 
  arrange(desc(n))
```
There are `r n_distinct(instacart$aisle)` aisles. The most item are ordered from fresh vegetables aisles. 

Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered

```{r}
instacart |> 
  count(aisle) |> 
  filter(n > 10000) |> 
  mutate(aisle = fct_reorder(aisle, n)) |> 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

```

Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r}
instacart |>
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) |>
  group_by(aisle) |>
  count(product_name) |> 
  mutate(rank = min_rank(desc(n))) |> 
  filter(rank < 4) |> 
  arrange(desc(n)) |>
  knitr::kable()  
```

Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week.

```{r}
instacart |>
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) |>
  group_by(product_name, order_dow) |>
  summarize(mean_hour = mean(order_hour_of_day)) |>
  pivot_wider(
    names_from = order_dow, 
    values_from = mean_hour) |>
  knitr::kable(digits = 2)
```

# Problem 2 (BRFSS)

Load the data data from the p8105.datasets package using:
```{r}
data("brfss_smart2010")
brfss=brfss_smart2010
```
 
*Data cleaning:

*format the data to use appropriate variable names;
*focus on the “Overall Health” topic
*include only responses from “Excellent” to “Poor”
*organize responses as a factor taking levels ordered from “Poor” to “Excellent”

Specify response column (show all the values):
```{r}
brfss=brfss |>
  janitor::clean_names() |>
  mutate(data_value=as.double(data_value))
unique(brfss$response)
```
The responses include "Excellent", "Very Good", "Good", "Fair", and "Poor".

Organize responses as a factor taking levels ordered from "Poor" to "Excellent":
```{r}
brfss=brfss |>
  filter(topic=="Overall Health",
    response %in% c("Poor", "Fair", "Good", "Very Good", "Excellent")) |>
  mutate(response=factor(response, levels=c("Poor", "Fair", "Good", "Very Good", "Excellent")))
```
Therefore, the order now is "Poor", "Fair", "Good", "Very Good", and "Excellent".

In 2002, states were observed at 7 or more locations
```{r}
brfss |>
  filter(year==2002) |>
  group_by(locationabbr) |>
  summarize(n_stats = n_distinct(locationdesc)) |>
  filter(n_stats>=7)
```
CT, FL, MA, NC, NJ, and PA are states observed at 7 or more locations in 2002.

In 2010, states were observed at 7 or more locations
```{r}
brfss |>
  filter(year==2010) |>
  group_by(locationabbr) |>
  summarize(n_stats = n_distinct(locationdesc)) |>
  filter(n_stats>=7)
```
CA, CO, FL, MA, MD, NC, NE, NJ, NY, OH, PA, SC, TX, and WA are states observed at 7 or more locations in 2010.
